---
title: "Coancestry Calculator"
author: "Chris Simoes"
date: "3/4/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE)
knitr::opts_chunk$set(message=FALSE)
knitr::opts_chunk$set(cache=FALSE)
```

## Coancestry Calculator

This is a tool to calculate the coancestry index. You will need this:

1 - Import a .csv file with these 3 columns: ID, P1, P2 
ID - is the individual name or identification
P1 - is the parental 1 or mother
P2 - is the parental 2 or father
If the individual doesn't have P1 and P2, fill with 0 (zero).

Following there is an example with 3 possible parents and 2 half sib:

ID,P1,P2<br>
A,0,0<br>
B,0,0<br>
C,0,0<br>
D,A,B<br>
E,C,B<br>



```{r importingFile, echo = TRUE}

#define your work directory
setwd("~/Documents/Cornell/PLBRG7170")

#Importing the csv file
infile <- read.csv("pedigree.csv", header = T, sep = ",")

#Getting all IDs 
names <- infile$ID

#Preparing a matrix nxn with all IDs
coancestry <- matrix(data = "NA", nrow = nrow(infile), ncol = nrow(infile))
rownames(coancestry)<-names
colnames(coancestry)<-names

```

## Calculating the coancestry 

1. Number individuals so that parents precede their ospring (e.g., from oldest to youngest). This numbering ensures that, when calculating 
fUV = 1/2 (fDV + fSV ), V is not a descendant of U.

2. Founders are assumed not to have common ancestors. Therefore, for founders enter 0 on the o-diagonal and 1/2 on the diagonal

3. Non-founder i has parents d and s (d < i and s < i), and is represented by row i of the matrix. In row i, calculate elements 1 to i - 1 by
averaging the corresponding elements from rows d and s, following the rule fij = 1/2 (fdj + fsj)

4. Set element i of row i (i.e., the diagonal element) to 1/2(1+fsd), following
the rule for the coancestry of an individual with itself fii = 1/2(1 + fsd)

5. Complete column i by symmetry with row i

```{r calculating, echo=TRUE}
for(i in 1:nrow(infile)){
  if(infile$P1[i] == 0 & infile$P2[i] == 0){
    coancestry[i,i] <- 0.5
    j=1
    while(j<i){
      coancestry[i,j] <- 0
      coancestry[j,i] <- 0
      j=j+1
    }
  }else{
    index_1 = i
    # print(paste(infile[index_1,1],infile[index_1,2]))
    z=1
    for(z in 1:i){
      if(names[z] == infile$P1[index_1]){
        index_2 = z
      }else if (names[z] == infile$P2[index_1]){
        index_3 = z
      }
      z=z+1
    }
    j=1
    while(j<i){
      coancestry[i,j] <- round((as.numeric(coancestry[index_2,j])+as.numeric(coancestry[index_3,j]))/2,3)
      coancestry[j,i] <- coancestry[i,j]
      j=j+1
    }
    if(i==j){
    coancestry[i,j]<-round((0.5*(1+as.numeric(coancestry[index_2,index_3]))),3)
    }
  }
}
# library(dplyr)
final_table <- knitr::kable(coancestry)
final_table
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
